---
title: "Plotting Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plotting_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(ERK)
library(ggplot2)
library(gtable)
library(ggplotify)
library(grid)
library(showtext)
library(lemon)

```

## Set-up

First set up some sample data and plot with facets: 

```{r}

dsamp <- ggplot2::diamonds[sample(nrow(diamonds), 1000), ]

```

```{r, fig.width = 8, fig.height =6}

d <- ggplot(dsamp, aes(carat, price)) +
 geom_point(aes(colour = clarity)) +
 facet_grid(color~cut)

d 
```

## theme_ERK()

This custom theme is based on theme_classic() and removes borders around the strip labels as well as some other minor changes. Use facets=TRUE when faceting in x and y directions to remove extra space between facets. If only faceting in one direction facets=FALSE will probably produce better results. 

```{r, fig.width = 8, fig.height =6}

d <- ggplot(dsamp, aes(carat, price)) +
  theme_ERK(facets=TRUE) +
 geom_point(aes(colour = clarity)) +
   facet_rep_grid(color~cut) 
d 

```
## overall_facet_labels

Add overall facet labels on both the top and right. 

```{r, fig.width = 8, fig.height =6}

overall_facet_labels(p = d, labelR = "Right Label", labelT = "Top Label", where = "both")

```

Only facet horizontally and add overall facet label at the top: 

```{r, fig.width = 8, fig.height =4}

d2 <- ggplot(dsamp, aes(carat, price)) +
  theme_ERK(facets=FALSE) +
 geom_point(aes(colour = clarity)) +
 facet_grid(.~cut)

overall_facet_labels(d2, labelR = "Right Label", labelT = "Top Label", where = "top")
```

Only facet vertically and add overall facet label at the right: 


```{r, fig.width = 5, fig.height = 6}

d2 <- ggplot(dsamp, aes(carat, price)) +
  theme_ERK(facets=FALSE) +
 geom_point(aes(colour = clarity)) +
 facet_grid(cut~.)

overall_facet_labels(p = d2, labelR = "Right Label", labelT = "Top Label", where = "right")

```

## using showtext package

Note: I have had issues getting symbols (e.g., minus sign) and subscripts to show up correctly with Computer Modern Font. Recommend use of extrafonts package instead in these cases.  

```{r, fig.showtext = TRUE, fig.width = 8, fig.height =6}


font_add(family = "Computer Modern", regular = "/usr/share/fonts/truetype/computer_modern/CMU_Serif_Roman.ttf")
# font_families()

showtext_auto()
 

d3 <- d + theme(text = element_text(family = "Computer Modern"))

d4 <- overall_facet_labels(p = d3, labelR = "Right Label", labelT = "Top Label", where = "both")

d4

```

Here we also use reposition_legend from the lemon package: 

```{r, fig.showtext = TRUE, fig.width = 7}


# 
# ## silly wrapper to fool ggplot2
# #' @export
# border_custom <- function(...){
#   structure(
#     list(...), # this ... information is not used, btw
#     class = c("element_custom","element_blank", "element") # inheritance test workaround
#   )
# 
# }
# 
# # these parts ensure only the left and bottom borders of each facet are there
# #' @export
# element_grob.element_custom <- function(element, ...)  {
#   #https://www.howtobuildsoftware.com/index.php/how-do/b1YQ/r-plot-ggplot2-r-grid-removing-right-border-from-ggplot2-graph
# 
#   segmentsGrob(x0 = c(1,0),
#                y0 = c(0,0),
#                x1 = c(0,0),
#                y1 = c(0,1), gp=gpar(lwd=2))
# }
# 
# 
# get_legend<-function(myggplot){
#   tmp <- ggplot_gtable(ggplot_build(myggplot))
#   leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
#   legend <- tmp$grobs[[leg]]
#   return(legend)
# }

showtext_auto()
 
d2 <- ggplot(dsamp, aes(carat, price)) +
 geom_point(aes(colour = clarity)) +
 facet_grid(.~cut)

d2 <- d2 + theme(text = element_text(family = "Computer Modern"))

d3 <- reposition_legend(d2 , 'center',  panel = "panel-1-1")

overall_facet_labels(d3, labelR = "Right Label", labelT = "Top Label", where = "top", input = "grob")

```

## using extrafonts package

```{r}

# devtools::install_github("wch/extrafont")
# install.packages("extrafontdb")
# library(ggplot2)
# library(extrafont)
# 
# font_import()
# 
# library(fontcm)
# font_install('fontcm')
# 
# fonts()
# # View(fonttable())
# 
# loadfonts()
# 
# pdf("ggplot_cm.pdf", width=4, height=4)
# p <- qplot(c(1,5), c(1,5)) +
#   xlab("Made with CM fonts") + ylab("Made with CM fonts") +
#   ggtitle("Made with CM fonts")
# 
# # Equation
# eq <- "italic(sum(frac(1, n*'!'), n==0, infinity) ==
#        lim(bgroup('(', 1 + frac(1, n), ')')^n, n %->% infinity))"
# 
# # Without the new fonts
# p + annotate("text", x=3, y=3, parse=TRUE, label=eq)
# 
# # With the new fonts
# p + annotate("text", x=3, y=3, parse=TRUE, family="CM Roman", label=eq) +
#   theme(text         = element_text(size=16, family="CM Roman"),
#         axis.title.x = element_text(face="italic"),
#         axis.title.y = element_text(face="bold"))
# 
# dev.off()
# 
# # Embed the fonts
# embed_fonts("ggplot_cm.pdf", outfile="ggplot_cm_embed.pdf")
```









